Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL

### 2020-12-07 main.tex
- paper changes

	\subsubsection{Normalized advantage function}

@@

	\subsubsection{Overview of the used algorithms}
	\Cref{tab:overview_algorithms} shows a comparison of the discussed algorithms. Details are provided in the appendix. All algorithms are suitable for many control tasks as common in accelerator operation. Their strength lies in the high sample-efficiency, but they differ in representational power and noise robustness. \\
	In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.
\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
	\caption{\label{tab:overview_algorithms}%
		An overview of the algorithms discussed in this paper.
	}
	\begin{ruledtabular}
		\begin{tabular}{lllll}
			\textrm{Algorithm}&
			\textrm{Class}&
			\textrm{ \begin{tabular}{@{}c@{}}Noise\\robustness\end{tabular}}&
			\textrm{ \begin{tabular}{@{}c@{}}Represent.\\power\end{tabular}}&
			\textrm{ \begin{tabular}{@{}c@{}}Sample\\efficiency\end{tabular}}\\
			\colrule
			\emph{NAF} & model-free & low &low &high\\
			\emph{NAF2} & model-free & high &low& high\\
			\emph{ME-TRPO} & model-based & low& high& high\\
			\emph{AE-DYNA} & model-based & high& high& high\\
		\end{tabular}
	\end{ruledtabular}
\end{table}
\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
	\caption{\label{tab:overview_algorithms_results}%
		An overview over the experimental results of the different algorithms.
	}
\begin{ruledtabular}
\begin{tabular}{llrr}
%\toprule
&    &  mean &   std \\
%\midrule
ae\_dyna & cum & -1.44 &  1.10 \\
me\_trpo & cum & -1.95 &  1.76 \\
naf\_double & cum & -0.58 &  0.72 \\
naf\_single & cum & -0.66 &  0.80 \\
ae\_dyna & final &  0.04 &  0.06 \\
me\_trpo & final &  0.01 &  0.04 \\
naf\_double & final & -0.00 &  0.02 \\
naf\_single & final &  0.00 &  0.02 \\
ae\_dyna & nr &  4.28 &  1.26 \\
me\_trpo & nr &  5.46 &  2.32 \\
naf\_double & nr &  2.67 &  1.33 \\
naf\_single & nr &  2.56 &  1.26 \\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

@@

	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps. In the MFRL, 700 data-points were used and in the MBRL 450-500 data-points were used. In all experiments the 50 verification episodes were finished successfully.\\

@@

	\subsection{NAF2 details}\label{appendix:naf2}
	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network and in all other cases a single network is used. 
	 The term \emph{smoothing} indicates that a small clipped noise is added on the actions to stabilize the network training as:
	\begin{equation}
		\ba(\bs) = \text{clip}\left(\mu_{\theta_{\text{targ}}}(\bs) + \text{clip}(\epsilon,-c,c), \ba_{Low}, \ba_{High}\right),
	\end{equation}
	where $\epsilon \sim \mathcal{N}(0, \sigma)$. $c>0$ denotes the clipping coefficient and $ \ba_{Low}, \ba_{High}$ the minimal and maximal possible action. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
	The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
	\begin{align}
		y(\reward_t,\bs_{t+1}, d_t) = \reward_t + \gamma (1 - d_t) \min_{i=1,2} V_{\theta_{i, \text{targ}}}(\bs_{t+1}),
	\end{align}
	with $d$ is one if the episode is finished and 0 otherwise.
	Then both are learned by regressing to this target (using the tuples from the data buffer $\mathcal D$ ):
	\begin{multline}
		L(\theta_i, {\mathcal D}) \\= \mathbb E_{(\bs_{t},\ba_{t},r_t,\bs_{t+1},d_t) \sim {\mathcal D}}{
			\Bigg( Q_{\theta_i}(\bs_{t},\ba_{t}) - y(r_t,\bs_{t+1},d_t) \Bigg)^2
		},
	\\
	i \in \{1,2\}
	\end{multline}
	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}(\bs)$.\\
	The results are shown in \cref{fig:comparsion_smoothing_small}. One sees the cumulative reward per episode for a training of a total of 100 episodes. As mentioned, the curve labelled \emph{clipping} corresponds to the double network including \emph{smoothing} and shows the best overall stability during the training yielding quickly a high reward. Also the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the sightly decreased stability. The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
%		\caption{Cumulative reward of different NAF implementations as discussed in the text.}
%		\label{fig:comparsion_smoothing}
%	\end{figure}
	\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_naf}
	\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text on the \emph{inverted pendulum} without noise.}
	\label{fig:comparsion_smoothing_small}
\end{figure}

@@

%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
%	One issue of MBRL methods is asymptotic performance. Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL. We call it the \emph{reality gap}. There are ideas to attack the problem, e.g. by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
%		\caption{The training .}
%		\label{fig:Learning_evolution}
%	\end{figure}
\section{Still open}
	\begin{itemize}
	\item Reference and code clean up.
	\item Conclusion and outlook\dots
	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
	\item Check the plots, which might be already okay!
%	\item Noise benchmarks with pendulum - done
%	\item AE details from papers
%	\item Benchmark NAF2 with pendulum - done
%	\item Compare NAF2 and AE-DYNA
%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
%	\item Tailored to accelerators - short horizons are standard?
%	\item Long waiting time in MBRL	
%	\item Vary the model size
%	\item Noise sensitivity as mentioned in Benchmark paper
%	\item Improvements of the algorithms:
%	\begin{itemize}
%		\item Adding the noise feature
%		\item Jumping feature using \emph{SAC}
%		\item Noise in the policy
%		\item discuss different approaches to, e.g. dynamic waiting time
%		\item usually prior not discussed in papers
%	\end{itemize}
\end{itemize}

>>>>

	\subsubsection{Normalized advantage function}\label{ss:Normalized advantage function}

@@

	\subsubsection{Overview of the used algorithms}
	\Cref{tab:overview_algorithms} shows a comparison of the discussed algorithms. Details are provided in the appendix. All algorithms are suitable for many control tasks as common in accelerator operation. Their strength lies in the high sample-efficiency, but they differ in representational power and noise robustness. \\
	In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.
\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
	\caption{\label{tab:overview_algorithms}%
		An overview of the algorithms discussed in this paper.
	}
	\begin{ruledtabular}
		\begin{tabular}{lllll}
			\textrm{Algorithm}&
			\textrm{Class}&
			\textrm{ \begin{tabular}{@{}c@{}}Noise\\robustness\end{tabular}}&
			\textrm{ \begin{tabular}{@{}c@{}}Represent.\\power\end{tabular}}&
			\textrm{ \begin{tabular}{@{}c@{}}Sample\\efficiency\end{tabular}}\\
			\colrule
			\emph{NAF} & MFRL & low &low &high\\
			\emph{NAF2} &MFRL & high &low& high\\
			\emph{ME-TRPO} & MBRL & low& high& high\\
			\emph{AE-DYNA} & MBRL & high& high& high\\
		\end{tabular}
	\end{ruledtabular}
\end{table}

@@

	\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
		\caption{\label{tab:overview_verification}%
			An overview over the verification results of the different algorithms.
		}
		\begin{ruledtabular}
			\begin{tabular}{lrrrr}
				%\toprule
				&  data&  nr&  cum &   final \\
				%\midrule
				\emph{AE-DYNA} &500 &  4.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
				\emph{ME-TRPO} &450&  5.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
				\emph{NAF} &1 074&  2.67 $\pm$  1.33 & -0.58 $\pm$  0.72 &0.00 $\pm$  0.02\\
				\emph{NAF2} &824& 2.56 $\pm$  1.26 & -0.66 $\pm$  0.80 &0.00 $\pm$  0.02\\
				%ae\_dyna & final &  0.04 &  0.06 \\
				%me\_trpo & final &  0.01 &  0.04 \\
				%naf\_double & final & -0.00 &  0.02 \\
				%naf\_single & final &  0.00 &  0.02 \\
				%ae\_dyna & nr &  4.28 &  1.26 \\
				%me\_trpo & nr &  5.46 &  2.32 \\
				%naf\_double & nr &  2.67 &  1.33 \\
				%naf\_single & nr &  2.56 &  1.26 \\
			\end{tabular}
		\end{ruledtabular}
	\end{table}

@@

	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps. The results of the verification is summarised in \cref{tab:overview_verification}. In the MFRL, around 800-1000 and in the MBRL 450-500 data-points were used. In all experiments the 50 verification episodes were finished successfully. \emph{NAF} performs better in terms of episode length while the average reward in the \emph{DYNA}-style algorithms is higher. The experiments were done on different days, hence under slightly different conditions.\\

@@

	\subsection{NAF2 details}\label{appendix:naf2}
	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network, as introduced in \cref{ss:Normalized advantage function} and in all other cases a single network is used. 
	 The term \emph{smoothing} indicates that a small clipped noise is added on the actions to stabilize the network training as:
	\begin{equation}
		\ba(\bs) = \text{clip}\left(\mu_{\theta_{\text{targ}}}(\bs) + \text{clip}(\epsilon,-c,c), \ba_{Low}, \ba_{High}\right),
	\end{equation}
	where $\epsilon \sim \mathcal{N}(0, \sigma)$. $c>0$ denotes the clipping coefficient and $ \ba_{Low}, \ba_{High}$ the minimal and maximal possible action. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
	The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
	\begin{align}
		y(\reward_t,\bs_{t+1}, d_t) = \reward_t + \gamma (1 - d_t) \min_{i=1,2} V_{\theta_{i, \text{targ}}}(\bs_{t+1}),
	\end{align}
	with $d$ is one if the episode is finished and 0 otherwise.
	Then both are learned by regressing to this target (using the tuples from the data buffer $\mathcal D$ ):
	\begin{multline}
		L(\theta_i, {\mathcal D}) \\= \mathbb E_{(\bs_{t},\ba_{t},r_t,\bs_{t+1},d_t) \sim {\mathcal D}}{
			\Bigg( Q_{\theta_i}(\bs_{t},\ba_{t}) - y(r_t,\bs_{t+1},d_t) \Bigg)^2
		},
	\\
	i \in \{1,2\}
	\end{multline}
	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}(\bs)$.\\
	The results are shown in \cref{fig:comparsion_smoothing_small}. One sees the cumulative reward per episode for a training of a total of 100 episodes. As mentioned, the curve labelled \emph{clipping} corresponds to the double network including \emph{smoothing} and shows the best overall stability during the training yielding quickly a high reward. Also the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the sightly decreased stability. The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
%		\caption{Cumulative reward of different NAF implementations as discussed in the text.}
%		\label{fig:comparsion_smoothing}
%	\end{figure}
	\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_naf}
	\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text on the \emph{inverted pendulum} without noise.}
	\label{fig:comparsion_smoothing_small}
\end{figure}

@@

%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
%	One issue of MBRL methods is asymptotic performance. Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL. We call it the \emph{reality gap}. There are ideas to attack the problem, e.g. by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
%		\caption{The training .}
%		\label{fig:Learning_evolution}
%	\end{figure}
\section{Still open}
	\begin{itemize}
	\item Reference and code clean up.
	\item Conclusion and outlook\dots
	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
	\item Add the noise policy of NAF
	\item Check the plots, which might be already okay!
%	\item Noise benchmarks with pendulum - done
%	\item AE details from papers
%	\item Benchmark NAF2 with pendulum - done
%	\item Compare NAF2 and AE-DYNA
%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
%	\item Tailored to accelerators - short horizons are standard?
%	\item Long waiting time in MBRL	
%	\item Vary the model size
%	\item Noise sensitivity as mentioned in Benchmark paper
%	\item Improvements of the algorithms:
%	\begin{itemize}
%		\item Adding the noise feature
%		\item Jumping feature using \emph{SAC}
%		\item Noise in the policy
%		\item discuss different approaches to, e.g. dynamic waiting time
%		\item usually prior not discussed in papers
%	\end{itemize}
\end{itemize}
